---
title: 'Logistic Regression and KNN Classifiers'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r, include=FALSE}
library(tidyverse)
# library(ggformula)
# library(gridExtra)
library(ISLR2)
```

<style>
  .col2 {
    columns: 2 200px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 200px; /* chrome, safari */
    -moz-columns: 2 200px;    /* firefox */
  }
  .col3 {
    columns: 3 100px;
    -webkit-columns: 3 100px;
    -moz-columns: 3 100px;
  }
</style>



## K-Nearest Neighbors Regression

* Non-parametric approach

* Given a value for $K$ and a test data point $x_0$,

$$\hat{f}(x_0)=\dfrac{1}{K} \sum_{x_i \in \mathcal{N}_0} y_i=\text{Average} \ \left(y_i \ \text{for all} \ i:\ x_i \in \mathcal{N}_0\right) $$

where $\mathcal{N}_0$ is known as the **neighborhood** of $x_0$.


* The method is based on the concept of closeness of $x_i$'s from $x_0$ for inclusion in the neighborhood $\mathcal{N}_0$. Usually, the **Euclidean distance** is used as a measure of closeness. The Euclidean distance between two $p$-dimensional vectors $\mathbf{a}=(a_1, a_2, \ldots, a_p)$ and $\mathbf{b}=(b_1, b_2, \ldots, b_p)$ is

$$||\mathbf{a}-\mathbf{b}||_2 = \sqrt{(a_1-b_1)^2 + (a_2-b_2)^2 + \ldots + (a_p-b_p)^2}$$


<!-- ## K-Nearest Neighbors Regression (single predictor): Fit -->

<!-- **Restaurant Outlets Profit dataset** -->

<!-- ```{r, message=FALSE} -->
<!-- library(caret)   # load the caret package -->
<!-- ``` -->

<!-- * **1-NN regression** -->

<!-- ```{r} -->
<!-- knnfit1 <- knnreg(profit ~ population, data = outlets, k = 1)   # 1-nn regression -->
<!-- ``` -->

<!-- ```{r} -->
<!-- predict(knnfit1, newdata = data.frame(population = 6))  # 1-nn prediction -->
<!-- ``` -->


<!-- * **5-NN regression** -->

<!-- ```{r} -->
<!-- knnfit5 <- knnreg(profit ~ population, data = outlets, k = 5)   # 5-nn regression -->
<!-- ``` -->

<!-- ```{r} -->
<!-- predict(knnfit5, newdata = data.frame(population = 6))  # 5-nn prediction -->
<!-- ``` -->

<!-- ## K-Nearest Neighbors Regression (single predictor): Prediction -->

<!-- **Ames Housing dataset** -->

<!-- ```{r} -->
<!-- nearest_neighbors <- ames %>% -->
<!--   select(Sale_Price, Gr_Liv_Area) %>% -->
<!--   mutate(distance = sqrt((1008-Gr_Liv_Area)^2)) %>%   # calculate distance -->
<!--   arrange(distance)   # sort by increasing distance -->
<!-- ``` -->

<!-- ```{r} -->
<!-- predict(knnfit1, newdata = data.frame(population = 1008))  # 1-nn prediction -->
<!-- ``` -->

<!-- ```{r} -->
<!-- predict(knnfit5, newdata = data.frame(population = 1008))  # 5-nn prediction -->
<!-- ``` -->


<!-- ## Regression Methods: Comparison -->

<!-- **Restaurant Outlets Profit dataset** -->

<!-- ```{r, echo=FALSE, fig.align='center', fig.height=5, fig.width=8, fig.cap="dashed cyan: 1-nn fit, dotted red: 5-nn fit, blue: linear regression fit"} -->
<!-- slrfit <- lm(profit ~ population, data = outlets)   # fit the SLR model -->

<!-- pop_seq <- seq(min(outlets$population, na.rm = TRUE), max(outlets$population, na.rm = TRUE), 0.01) -->

<!-- # obtain predictions for all training data points -->
<!-- # knn_1 <- predict(knnfit1, newdata = data.frame(population = min(outlets$population, na.rm = TRUE):max(outlets$population, na.rm = TRUE))) -->
<!-- knn_1 <- predict(knnfit1, newdata = data.frame(population = pop_seq)) -->
<!-- # knn_5 <- predict(knnfit5, newdata = data.frame(population = min(outlets$population, na.rm = TRUE):max(outlets$population, na.rm = TRUE))) -->
<!-- knn_5 <- predict(knnfit5, newdata = data.frame(population = pop_seq)) -->
<!-- # p <-  predict(slrfit, newdata = data.frame(population = min(outlets$population, na.rm = TRUE):max(outlets$population, na.rm = TRUE))) -->
<!-- p <-  predict(slrfit, newdata = data.frame(population = pop_seq)) -->

<!-- # column bind original data with predicted values -->
<!-- # ames1 <- outlets %>% select(profit, population) %>% filter(!is.na(population)) -->
<!-- # predictions <- data.frame(population = min(outlets$population, na.rm = TRUE):max(outlets$population, na.rm = TRUE), -->
<!--                           # linear = p, knn_1, knn_5) -->
<!-- predictions <- data.frame(population = pop_seq, -->
<!--                           linear = p, knn_1, knn_5) -->

<!-- # plot the three models -->
<!-- ggplot(data = outlets, aes(x = population, y = profit)) + -->
<!--   geom_point() + -->
<!--   geom_line(data = predictions, aes(x = population, y = knn_1), color = "cyan", linetype = "dashed", linewidth = 1) +   # 1-nn regression -->
<!--   geom_line(data = predictions, aes(x = population, y = knn_5), color = "red", linetype = "dotted", linewidth = 1) +   # 5-nn regression -->
<!--   geom_line(data = predictions, aes(x = population, y = linear), color = "blue", linewidth = 1) + -->
<!--   theme_bw() -->

<!-- ``` -->


<!-- ## <span style="color:blue">Your Turn!!!</span> -->

<!-- For the **Advertising** dataset, create a 10-nearest neighbors fit **knnfit10** with **sales** as response and **TV** as predictor. Obtain the predicted **sales** for $\text{TV} = 225,000$. -->


<!-- ## <span style="color:blue">Question!!!</span> -->

<!-- As $k$ in KNN regression increases, -->

<!-- * the flexibility of the fit $\underline{\hspace{5cm}}$ (increases/decreases) -->

<!-- * the bias of the fit $\underline{\hspace{5cm}}$ (increases/decreases) -->

<!-- * the variance of the fit $\underline{\hspace{5cm}}$ (increases/decreases) -->


## K-Nearest Neighbors Regression (multiple predictors)

It is important to **scale (subtract mean and divide by standard deviation)** the predictors when considering KNN regression so that the Euclidean distance is not dominated by a few of them with large values.

<!-- **House Prices dataset** -->

<!-- ```{r} -->

<!-- # scale predictors -->
<!-- house_prices_scaled <- data.frame(size_scaled = scale(house_prices$size), -->
<!--                                   num_bedrooms_scaled = scale(house_prices$num_bedrooms), -->
<!--                                   price = house_prices$price) -->

<!-- head(house_prices_scaled)   # first six observations -->

<!-- ``` -->


<!-- ## K-Nearest Neighbors Regression (multiple predictors) -->

<!-- **House Prices dataset** -->

<!-- ```{r, message=FALSE} -->
<!-- library(caret)   # load library -->
<!-- ``` -->

<!-- ```{r} -->

<!-- knnfit10 <- knnreg(price ~ size_scaled + num_bedrooms_scaled, data = house_prices_scaled, k = 10)   # 10-nn regression -->

<!-- ``` -->


<!-- It is also important to apply scaling to test data points before prediction. Suppose, you want predictions for `size` = 2000 square feet, and `num_bedrooms` = 3, then -->

<!-- ```{r} -->
<!-- # obtain 10-nn prediction -->

<!-- predict(knnfit10, newdata = data.frame(size_scaled = (2000 - mean(house_prices$size, na.rm = TRUE))/sd(house_prices$size, na.rm = TRUE), -->
<!--                                      num_bedrooms_scaled = (3 - mean(house_prices$num_bedrooms))/sd(house_prices$num_bedrooms))) -->
<!-- ``` -->


## Linear Regression vs K-Nearest Neighbors


<!-- * Linear regression is a parametric approach (with restrictive assumptions), KNN is non-parametric. -->

<!-- * Linear regression works for regression problems ($Y$ numerical), KNN can be used for both regression and classification ($Y$ qualitative). -->

<!-- * Linear regression is interpretable, KNN is not. -->

<!-- * Linear regression can accommodate qualitative predictors and can be extended to include interaction terms as well. Using Euclidean distance with KNN does not allow for qualitative predictors. -->

<!-- * In terms of prediction, KNN can be pretty good for small $p$, that is, $p \le 4$ and large $n$. Performance of KNN deteriorates as $p$ increases - curse of dimensionality. -->


<center>
|  |  Linear Regression | KNN Regression |
|-----------|----------|-----------
| Type |  |  |
| Task |  |  |
| Predictive Performance |  |  |
| Interpretability |  |  |
| Other |  |  |

</center>
\




<!-- ## Linear Regression vs K-Nearest Neighbors (Recap) -->


<!-- * Linear regression is a parametric approach (with restrictive assumptions), KNN is non-parametric. -->

<!-- * Linear regression works for regression problems ($Y$ numerical), KNN can be used for both regression and classification ($Y$ qualitative). -->

<!-- * Linear regression is interpretable, KNN is not. -->

<!-- * Linear regression can accommodate qualitative predictors and can be extended to include interaction terms as well. Using Euclidean distance with KNN does not allow for qualitative predictors. -->

<!-- * In terms of prediction, KNN can be pretty good for small $p$, that is, $p \le 4$ and large $n$. Performance of KNN deteriorates as $p$ increases - curse of dimensionality. -->


<!-- <center> -->
<!-- |  |  Linear Regression | KNN Regression | -->
<!-- |-----------|----------|----------- -->
<!-- | Type |  |  | -->
<!-- | Task |  |  | -->
<!-- | Predictive Performance |  |  | -->
<!-- | Interpretability |  |  | -->
<!-- | Other |  |  | -->

<!-- </center> -->
<!-- \ -->




## Classification Problems

* Response $Y$ is qualitative (categorical).

* The objective is to build a classifier $\hat{Y}=\hat{C}(\mathbf{X})$ that assigns a class label to a future unlabeled (unseen) observation and understand the relationship between the predictors and response.

* There can be two types of predictions based on the research problem.

  + Class probabilities
  
  + Class labels

<!-- Often we are more interested in estimating the probabilities than actual class labels. -->


<!-- ## Classification Problems: Example -->

<!-- ## Logistic Regression -->

<!-- * Supervised learning -->

<!-- * Classification (even though the term contains "Regression") -->

<!-- * Parametric approach -->

<!-- **Default dataset** -->

<!-- ```{r, message=FALSE} -->
<!-- library(ISLR2)   # load library -->
<!-- data("Default")   # load dataset -->
<!-- ``` -->

<!-- ```{r} -->
<!-- head(Default)   # print first six observations -->
<!-- ``` -->

<!-- ```{r,message=FALSE} -->
<!-- table(Default$default)   # class frequencies -->
<!-- ``` -->


<!-- **We will consider `default` as the response variable.** -->


<!-- ## Classification Problems: Example -->

<!-- For some algorithms, we might need to convert the categorical response to numeric (0/1) values. -->

<!-- **Default dataset** -->

<!-- ```{r,message=FALSE} -->
<!-- Default$default_id <- ifelse(Default$default == "Yes", 1, 0)   # create 0/1 variable -->

<!-- head(Default, 10)   # print first ten observations -->
<!-- ``` -->




## K-Nearest Neighbors Classifier

Given a value for $K$ and a test data point $x_0$,
$$P(Y=j | X=x_0)=\dfrac{1}{K} \sum_{x_i \in \mathcal{N}_0} I(y_i = j)$$

where $\mathcal{N}_0$ is known as the **neighborhood** of $x_0$.


For classification problems, the predictions are obtained in terms of **majority vote** (unlike in regression where predictions are obtained by averaging).

Let's implement this classifier in R.


<!-- ## K-Nearest Neighbors Classifier: Build Model  -->

<!-- **Default dataset** -->

<!-- response ($Y$): `default` and predictor ($X$): `balance` -->

<!-- ```{r} -->
<!-- library(caret)   # load package 'caret' -->

<!-- knnfit <- knn3(default ~ balance, data = Default, k = 10)   # fit 10-nn model -->
<!-- ``` -->


<!-- ## K-Nearest Neighbors Classifier: Predictions  -->

<!-- **Default dataset** -->

<!-- * One can directly obtain the class label predictions as below. -->

<!-- ```{r} -->
<!-- knn_class_preds_1 <- predict(knnfit, newdata = Default, type = "class")   # obtain default class label predictions -->
<!-- ``` -->


<!-- * Otherwise, one can first obtain predictions in terms of probabilities and then convert them into class label predictions based on a threshold. -->

<!-- ```{r} -->
<!-- knn_prob_preds <- predict(knnfit, newdata = Default, type = "prob")   # obtain predictions as probabilities -->
<!-- ``` -->


<!-- ```{r} -->
<!-- threshold <- 0.5   # set threshold -->

<!-- knn_class_preds_2 <- factor(ifelse(knn_prob_preds[,2] > threshold, "Yes", "No"))   # obtain predictions as class labels -->
<!-- ``` -->


<!-- ## K-Nearest Neighbors Classifier: Performance  {.smaller} -->

<!-- **Default dataset** -->

<!-- ```{r} -->
<!-- # create confusion matrix -->

<!-- # use the following code only when all predictions are from the same class -->
<!-- # levels(knn_class_preds_1) = c("No", "Yes")  -->

<!-- confusionMatrix(data = knn_class_preds_1, reference = Default$default, positive = "Yes")    -->
<!-- ``` -->





## Logistic Regression {.smaller}

A logistic regression classifier is a common ML model for binary classification problems. It assumes a functional form of the relationship between the response and the predictors.
\
\

Consider a one-dimensional two-class problem.

* Transform the linear model $\beta_0 + \beta_1 \ X$ so that the output is a probability.

* Use **logistic** or **sigmoid** function $$g(t)=\dfrac{e^t}{1+e^t} \ \ \ \text{for} \ t \in \mathcal{R}$$

* Suppose $p(X)=P(Y=1|X)$. Then, $$p(X)=g\left(\beta_0 + \beta_1 \ X\right)=\dfrac{e^{\beta_0 + \beta_1 \ X}}{1+e^{\beta_0 + \beta_1 \ X}}$$

* $e \approx 2.71828$ is a mathematical constant (Euler's number).

* $\log \left(\dfrac{p(X)}{1-p(X)}\right) = \beta_0 + \beta_1 \ X$


Let's implement this model in R.



## Why Not Linear Regression? 

**Default dataset**

```{r, echo=FALSE}
Default$default_id <- ifelse(Default$default == "Yes", 1, 0)
# lrfit <- lm(default_id ~ balance, data = Default)   # fit SLR
# summary(lrfit$fitted.values)   # summary of y_hats
```

```{r, echo=FALSE,  fig.align='center', fig.width=7, fig.height=4.5}
ggplot(data = Default, aes(x = balance, y = default_id)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + 
  scale_y_continuous(breaks = c(0, 0.5, 1)) + labs(y = "Default = Yes is 1, otherwise 0")
```

Linear regression does not model probabilities well. Linear regression might produce probabilities less than zero or bigger than one.


## Logistic Regression

**Default dataset**

```{r , fig.align='center', fig.width=7, fig.height=4.5, echo = FALSE}
logregfit <- glm(default ~ balance, data = Default, family = binomial)   # fit logistic regression model

ggplot(data = Default, aes(x = balance, y = ifelse(default == "Yes", 1, 0))) +
  geom_point() +
  geom_line(aes(y = logregfit$fitted.values), color = "blue") + 
  scale_y_continuous(breaks = c(0, 0.5, 1)) +
  labs(y = "Default = Yes is 1, otherwise 0")
```



## Why Not Linear Regression?

Suppose we have a response $Y$,

<center>
$Y=\begin{cases}
1; & \text{if stroke} \\
2; & \text{if drug overdose} \\
3; & \text{if epileptic seizure}
\end{cases}$
</center>
\

Linear regression suggests an ordering, and in fact implies that the difference between stroke and drug overdose is the same as between drug overdose and epileptic seizure.



<!-- ## Logistic Regression -->

## Logistic Regression: Predictions and Interpretation

**Default dataset**

For `balance`=\$700,

<!-- $$\hat{p}(X)=\dfrac{e^{\hat{\beta}_0+\hat{\beta}_1 X}}{1+e^{\hat{\beta}_0+\hat{\beta}_1 X}}=\dfrac{e^{-10.465328 + (0.005389 \times 700)}}{1+e^{-10.465328 + (0.005389 \times 700)}}=0.0012$$ -->

$$\hat{p}(X)=\dfrac{e^{b_0+b_1 X}}{1+e^{b_0+b_1 X}}=\dfrac{e^{-10.68 + (0.005516 \times 700)}}{1+e^{-10.68 + (0.005516 \times 700)}}=0.0011$$


<!-- ```{r, message=FALSE} -->
<!-- predict(logregfit, newdata = data.frame(balance = 700))   # obtain log-odds predictions -->
<!-- ``` -->

<!-- ```{r} -->
<!-- predict(logregfit, newdata = data.frame(balance = 700), type = "response")   # obtain probability predictions -->
<!-- ``` -->

**Interpretations**: A positive $\hat{\beta}_1$ indicates that a higher balance is associated with a larger log odds of default, in other words, a higher probability of default. However, the change in the probability of default due to a one-unit change in balance depends on the current balance value.



## Logistic Regression vs K-Nearest Neighbors


<!-- * Linear regression is a parametric approach (with restrictive assumptions), KNN is non-parametric. -->

<!-- * Linear regression works for regression problems ($Y$ numerical), KNN can be used for both regression and classification ($Y$ qualitative). -->

<!-- * Linear regression is interpretable, KNN is not. -->

<!-- * Linear regression can accommodate qualitative predictors and can be extended to include interaction terms as well. Using Euclidean distance with KNN does not allow for qualitative predictors. -->

<!-- * In terms of prediction, KNN can be pretty good for small $p$, that is, $p \le 4$ and large $n$. Performance of KNN deteriorates as $p$ increases - curse of dimensionality. -->


<center>
|  |  Logistic Regression | KNN Classifier |
|-----------|----------|-----------
| Type |  |  |
| Task |  |  |
| Predictive Performance |  |  |
| Interpretability |  |  |
| Other |  |  |

</center>
\


## Model Assessment: Confusion Matrix

<!-- <center> -->
<!-- |  | Predicted class labels | -->
<!-- |-----------|----------|----------- -->
<!-- |  |  0 | 1 | -->
<!-- |-----------|----------|----------- -->
<!-- | 0 | True Negative |  False Positive  | -->
<!-- | 1 | False Negative |  True Positive | -->
<!-- </center> -->
<!-- \ -->

<!-- <center> -->
<!-- |  | Reference class labels | -->
<!-- |-----------|----------|-----------| -->


<center>
|  |  Actual Positive/Event | Actual Negative/Non-event |
|-----------|----------|-----------
| Predicted Positive/Event |  |  |
| Predicted Negative/non-event |  |  |
</center>
\



## Model Assessment: ROC Curve

A Receiver Operating Characteristic curve, or ROC curve, is a graphical plot that illustrates the performance of a binary classifier model at varying threshold values.

The ROC curve with the `pROC` package in R plots sensitivity (recall) along the vertical y-axis and specificity along the horizontal x-axis.

The closer the curve hugs the upper left-hand corner, the better is the performance of the classifier. AUC refers to the Area Under this Curve.



